---
title: "R Crash Course"
output: html_notebook
---
by Jaehyuk CHOI

```{r}
browseURL("http://www.jaehyukchoi.net/phbs_en")
```

## Why use R?

* free (vs Matlab)
* open source (lots of volutarily developers distributing 'packages')
* statistics, scientific computation, graphics

## simple calculation

```{r}
3 + 4
```

```{r}
3 * 16
```

## Useful shortcut keys

* Ctrl-Enter, Ctrl-Enter on selection

* Ctrl-L: clear screen

* history, partical completion (Ctrl+Up)


## variable assignment

* variables are case-sensitive

```{r}
x <- 5
X <- 15
(x <- 5)
```

```{r}
y <- 8
x + y
```

## special constants
```{r}
TRUE 

T # same as TRUE

FALSE

F # same as FALSE

NA # Not Available, or Not Applicable

x.vec <- c(1,2)

Inf

1/0

NaN

0/0
```

## vector

```{r}
( x.vec <- 1:10 )  # matlab style a:b:c doesn't work

( x.vec <- c(1.3, 2.4, 3.5, 0.4) )

( y.vec <- c(5,4,3,2) )

x.vec + y.vec

x.vec * y.vec
```

## seq

```{r}
( x.vec <- seq(1, 5, 0.1) )

( x.vec <- seq(1, 100, 2) )

( x.vec <- seq(1, 5, by = 0.1) )

length(x.vec)

```

## rep function

```{r}
( y.vec <- rep(4, 10) )

( y.vec <- rep(c(3,4), 10) )

```

## indexing

```{r}
( x.vec <- 1:10 )

x.vec[4]

x.vec[1:5]

x.vec[-1]

x.vec[-(1:5)]

x.vec[ x.vec > 5 ]

x.vec[ x.vec > 5 ] <- 9
x.vec
```


## functions on vector
```{r}
( x.vec <- rnorm(10) )

sum( x.vec )

mean( x.vec )

var(x.vec)

prod( x.vec )

exp( x.vec )

log( x.vec )
```

## Probability distribution
```{r}
( x.vec <- seq(-2,2,by=0.1) )

d.vec <- dnorm( x.vec )

plot( x.vec, d.vec )

p.vec <- pnorm( x.vec )

plot( x.vec, p.vec )

( r.vec <- rnorm(10) )

runif(10)
dunif(c(0, 0.5, 1))
punif(c(0, 0.5, 1))
```


## Matrix
```{r}
( x.mat <- matrix(1:20) )

( x.mat <- matrix(1:20, 4) )

( x.mat <- matrix(1:20, nrow=5) )

( x.mat <- matrix(1:20, nrow=5, byrow=T) )
```

## Alternative ways of creating matrix
```{r}
( identity.mat <- diag( 10 ) )
( x.mat <- matrix( rnorm(25), ncol=5 ))
```

## functions on matrix
```{r}
exp( x.mat )

dim( x.mat )

diag(x.mat)
```

## matrix is also a vector (column-wise). 
* Rule-of-thumb is vector is a column vector by default

```{r}
length( x.mat )
x.mat
as.vector( x.mat )

```

## operations between vectors (or matrix) with different sizes

```{r}
( x.vec <- 1:20 )
( y.vec <- 1:4 ) 

x.vec * y.vec   # vector with smaller size is applied repeatedly.
y.vec * x.vec

( y.vec <- 1:3 ) 
x.vec * y.vec # error

( x.mat <- matrix( 1:20, nrow = 4 ) )
( y.vec <- 1:4 ) 

x.mat * y.vec  # the same applies to the matrix * vector multiplication
y.vec * x.mat
```

## matrix operations

```{r}

( x.mat <- matrix( 1:25, ncol=5 ) )
diag( x.mat )
t(x.mat)

x.mat[2:4,2:4]
x.mat[2:4,2:4] <- 0
x.mat

dim(x.mat)

( i.mat <- diag(5) )
( x.mat <- matrix(seq(1,25), ncol=5) )

i.mat * x.mat
i.mat %*% x.mat

( b.vec <- 1:5 )
( y.mat <- 10*diag(5) + matrix(rnorm(25), 5) )

( a.vec <- solve( y.mat, b.vec ) )   # Solve a such that Y * a = b
y.mat %*% a.vec - b.vec # confirm that Y * a = b

library("MASS")
( yinv.mat <- MASS::ginv(y.mat) )
yinv.mat %*% b.vec - a.vec # Inv(Y) * b = a

( eig <- eigen(y.mat) )
eig$values
eig$vectors

rowSums(x.mat)
rowMeans(x.mat)

colSums(x.mat)
colMeans(x.mat)

```

## Random number generation example: compute PI

We compute the area of the unit circle (PI) by random number generation (Monte-Carlo)

```{r}

xy <- matrix(runif(1e4), ncol=2)
head(xy)

th <- 2*pi*seq(0,1,by=0.01)
par(pty="s")
plot(xy, type="p", pch=".")
lines(cos(th),sin(th), col="red", lw=4)

xy <- matrix(runif(1e6), ncol=2)
sum( rowSums(xy*xy)<1 ) / nrow(xy) * 4

```



### Control sequence
```{r}

val <- 2
if ( val == 1 ) {
  1
} else if( val == 2) {
  2
} else {
  3
}
```


### For-loops, but avoid using it if possible
```{r}

i.sum <- 0
for(i in 1:10) {
  print(i)
  i.sum <- i.sum + i
} 
i.sum

( i.sum <- sum(1:10) )

```


## functions
```{r}

# Sourcing
source('crash_course_functions.R')

AddTwoNumbers( 2, 5 )

LogBase( 1000, 10 )

LogBase( 1000 )

LogBase( 1000, base = 2 )

```


## variable scope 

```{r}

# AddTwoNumbersWrong( 3, 4 ) # Error because yy is not defined in the function

yy <- 10

AddTwoNumbersWrong( 3, 4 ) # shouldn't work, but it works

```

## list
```{r}

a <- list(name="J Choi", phbs_number=12334)
a$name

```


## Dataframe (Car dataset)
```{r}

# install.packages("datasets")
head(mtcars)
#head(iris)

```

```{r}

c( ncol(mtcars), nrow(mtcars) )
colMeans(mtcars)
sapply(mtcars, sd)

## components by name or index
mtcars["Datsun 710", "hp" ]
mtcars[3,4]               

## column vector
mtcars[[4]]

mtcars[["hp"]]
mtcars$hp
mtcars[, "hp"]

## Column Slice
mtcars[2] 
mtcars["disp"]
mtcars[ c("disp", "drat")]

## Row slice
mtcars[ 2:5, ]
mtcars[ mtcars$mpg>30, ]
mtcars[ mtcars$mpg>30, ]$hp

```

## Linear regression
```{r}
library("datasets")
browseURL( "https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html" )
head(cars)

cars.fitted = lm(dist ~ speed, data = cars)
summary( cars.fitted )

plot( cars$speed, cars$dist, xlab = "Speed (mph)", ylab = "Stopping distance (ft)" )
abline(cars.fitted, col="red", lw=2)
```

```{r}
plot(cars, xlab = "Speed (mph)", ylab = "Stopping distance (ft)", las = 1, xlim = c(0, 25))
d <- seq(0, 25, length.out = 200)

for(degree in 1:4) {
  fm <- lm(dist ~ poly(speed, degree), data = cars)
  assign(paste("cars.fitted", degree, sep = "."), fm)
  lines(d, predict(fm, data.frame(speed = d)), col = degree)
}
anova(cars.fitted.1, cars.fitted.2, cars.fitted.3, cars.fitted.4)

```

## XL, CVS file
```{r}
wine <- read.csv("winequality-red.csv", header = TRUE, sep =";")
head(wine)
```
## Reading remote file on-line

```{r}
uci <- "http://archive.ics.uci.edu/ml/machine-learning-databases/"
uci <- paste(uci,"echocardiogram/echocardiogram.data",sep="")
ecc <- read.csv(uci)
head(ecc)
```

## Execution time
```{r}
i.sum <- 0;
system.time(
  for(i in 1:1e6) {
    i.sum <- i.sum + i
  } 
)

system.time( sum(as.numeric(1:1e6)) )
system.time( exp( rnorm(1e7) ) )
```

## avoid resizeing vector size
```{r}
( x.vec <- 1:10 )
length(x.vec)
x.vec[11] <- 11  ## no good !
length(x.vec)
```

## Google's R Style guide
```{r}
browseURL( 'https://cran.r-project.org/manuals.html' )
browseURL( 'https://google.github.io/styleguide/Rguide.xml' )
```

